---
title: "EVA Data Challenge"
date: "2023-04-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Q1

## Strategy

The general method for this question was to cluster the covariates, first sorting into seasons, and then clustering over the remaining dimensions (WindDirection and WindSpeed are converted into velocity components). An EM algorithm is used by mclust for maximum likelihood estimation. Initialisation of EM is performed using the partitions obtained from agglomerative hierarchical clustering. Using the resulting multivariate normal distributions, we are able to catagorise the data and fit a GPD to find the extreme quantile. In order to maximise the number of points in each cluster, we allow points in a cluster if they have a probability of over 0.25 of being in the cluster. We use a bootstrapping method to estimate central 50% confidence bounds. The clustering code and the bootstrapping code is included, as well as GPD analysis for the 10 clusters. It is possible to run the code but it is pretty slow and cluster allocations may vary slightly.




```{r}
library(mclust)
library(mvtnorm)
library(ismev)
library(ggplot2)
library(ggpubr)
library(stats)
library(corrplot)
library(reshape2)
library(dplyr)
library(evd)
AmaData <- read.csv('Amaurot.csv')
summary(AmaData)
```

```{r}
AmaData <- AmaData %>% 
  mutate(WindSpeedx = WindSpeed*cos(WindDirection))
AmaData <- AmaData %>% 
  mutate(WindSpeedy = WindSpeed*sin(WindDirection))
NewAmaData <- subset(AmaData, select = -c(Y,WindSpeed, WindDirection))
AmaDataS1 <- NewAmaData %>% filter(grepl('S1', Season))
AmaDataS2 <- NewAmaData %>% filter(grepl('S2', Season))
AmaDataS1 <- subset(AmaDataS1, select = -c(Season))
AmaDataS2 <- subset(AmaDataS2, select = -c(Season))
AmaDataS1_nona <- na.omit(AmaDataS1)
AmaDataS2_nona <- na.omit(AmaDataS2)
summary(AmaDataS1_nona)
summary(AmaDataS2_nona)


NewAmaDataYs <- subset(AmaData, select = -c(WindSpeed, WindDirection))
YsAmaDataS1 <- NewAmaDataYs %>% filter(grepl('S1', Season))
YsAmaDataS2 <- NewAmaDataYs %>% filter(grepl('S2', Season))
YsAmaDataS1 <- subset(YsAmaDataS1, select = -c(Season))
YsAmaDataS2 <- subset(YsAmaDataS2, select = -c(Season))
YsAmaDataS1_nona <- na.omit(YsAmaDataS1)
YsAmaDataS2_nona <- na.omit(YsAmaDataS2)
```

```{r}

loglikeS1 <- rep(0,9)
for (i in 2:10){
  num_clusters <- i
  gmmS1 <- Mclust(AmaDataS1_nona, G = num_clusters)
  loglikeS1[i-1] <- gmmS1$loglik
}

plot(seq(2, 10), -loglikeS1, pch = 19, col = "black", xlab = "Number Of Clusters", ylab = "Negative Log Likelihood")

# Add a red line
lines(seq(2, 10), -loglikeS1, col = "red")
```
```{r}
# Number of repetitions
num_repeats <- 10

# Initialize a matrix to store log likelihood values
loglike_matrix <- matrix(0, nrow = 9, ncol = num_repeats)

# Repeat the experiment
for (rep in 1:num_repeats) {
  for (i in 2:10) {
    num_clusters <- i
    gmmS1 <- Mclust(AmaDataS1_nona, G = num_clusters)
    loglike_matrix[i - 1, rep] <- -gmmS1$loglik
  }
}

# Calculate the average log likelihood values
avg_loglike <- rowMeans(loglike_matrix)

# Plot black points for the average values
plot(seq(2, 10), avg_loglike, pch = 19, col = "black", xlab = "Number Of Clusters", ylab = "Average Negative Log Likelihood")

# Add faint blue lines for individual experiments
for (rep in 1:num_repeats) {
  lines(seq(2, 10), loglike_matrix[, rep], col = "lightblue", lty = 2)
}

# Add a red line representing the average
lines(seq(2, 10), avg_loglike, col = "red")

```

```{r}

loglikeS2 <- rep(0,9)
for (i in 2:10){
  num_clusters <- i
  gmmS2 <- Mclust(AmaDataS2_nona, G = num_clusters)
  loglikeS2[i-1] <- gmmS2$loglik
}

plot(seq(2, 10), loglikeS2)
```

```{r}
#lets choose 5
num_clusters <- 5
gmmS1 <- Mclust(AmaDataS1_nona, G = num_clusters)
```

```{r}
#lets choose 5
num_clusters <- 5
gmmS2 <- Mclust(AmaDataS2_nona, G = num_clusters)
```


```{r}
S1clustparams <- gmmS1$parameters
alocsS1 <- gmmS1$z
num_points_S1 <- 9264

mixedprobsS1 <- rep(0,num_points_S1)

for (i in 1:num_points_S1){
  if (sum(alocsS1[i,]>0.25) >= 2){
    mixedprobsS1[i] <- 1
  }
  else{
    mixedprobsS1[i] <- 0
  }
}

```

```{r}
S2clustparams <- gmmS2$parameters
alocsS2 <- gmmS2$z
num_points_S2 <- 9281

mixedprobsS2 <- rep(0,num_points_S2)

for (i in 1:num_points_S2){
  if (sum(alocsS2[i,]>0.25) >= 2){
    mixedprobsS2[i] <- 1
  }
  else{
    mixedprobsS2[i] <- 0
  }
}

```


```{r}
cluster_alocsS1 <- list(
  clus1 = list(),
  clus2 = list(),
  clus3 = list(),
  clus4 = list(),
  clus5 = list()
)

for (i in 1:num_points_S1){
  for (j in 1:num_clusters){
    if (alocsS1[i,j]>0.25){
      cluster_alocsS1[[j]]<-append(cluster_alocsS1[[j]], i)
    }
  }
}
sum(mixedprobsS1)
```

```{r}
cluster_alocsS2 <- list(
  clus12 = list(),
  clus22 = list(),
  clus32 = list(),
  clus42 = list(),
  clus52 = list()
)

for (i in 1:num_points_S2){
  for (j in 1:num_clusters){
    if (alocsS2[i,j]>0.25){
      cluster_alocsS2[[j]]<-append(cluster_alocsS2[[j]], i)
    }
  }
}
sum(mixedprobsS2)
```




```{r}
cluster1_alocsS1 <- unlist(cluster_alocsS1$clus1, recursive = FALSE)
cluster2_alocsS1 <- unlist(cluster_alocsS1$clus2, recursive = FALSE)
cluster3_alocsS1 <- unlist(cluster_alocsS1$clus3, recursive = FALSE)
cluster4_alocsS1 <- unlist(cluster_alocsS1$clus4, recursive = FALSE)
cluster5_alocsS1 <- unlist(cluster_alocsS1$clus5, recursive = FALSE)
cluster1_alocsS2 <- unlist(cluster_alocsS2$clus12, recursive = FALSE)
cluster2_alocsS2 <- unlist(cluster_alocsS2$clus22, recursive = FALSE)
cluster3_alocsS2 <- unlist(cluster_alocsS2$clus32, recursive = FALSE)
cluster4_alocsS2 <- unlist(cluster_alocsS2$clus42, recursive = FALSE)
cluster5_alocsS2 <- unlist(cluster_alocsS2$clus52, recursive = FALSE)
```


```{r}
AmaTestData <- read.csv('AmaurotTestSet.csv')
AmaTestData <- AmaTestData %>% 
  mutate(WindSpeedx = WindSpeed*cos(WindDirection))
AmaTestData <- AmaTestData %>% 
  mutate(WindSpeedy = WindSpeed*sin(WindDirection))
NewAmaTestData <- subset(AmaTestData, select = -c(WindSpeed, WindDirection))

clusts <- data.frame(Season=character(),
                      Allocation=numeric(),
                      Probability=numeric())

for (i in 1:length(AmaTestData[,1])){
  vec1 <- NewAmaTestData[i,]
  if(vec1[5]== 'S1'){
    vec1 <- vec1[-5]
    densities <- rep(0,5)
    for (j in 1:5){
      densities[j] <- dmvnorm(vec1, mean = S1clustparams$mean[,j], sigma = S1clustparams$variance$sigma[,,j])
    }
    clust_probs <- densities/sum(densities)
    allocation <- which.max(clust_probs)
    probability <- clust_probs[allocation]
    row1 <- c('S1', allocation, probability)
    clusts <- rbind(clusts, row1)
  }
  if(vec1[5]== 'S2'){
    vec1 <- vec1[-5]
    densities <- rep(0,5)
    for (j in 1:5){
      densities[j] <- dmvnorm(vec1, mean = S2clustparams$mean[,j], sigma = S2clustparams$variance$sigma[,,j])
    }
    clust_probs <- densities/sum(densities)
    allocation <- which.max(clust_probs)
    probability <- clust_probs[allocation]
    row1 <- c('S2', allocation, probability)
    clusts <- rbind(clusts, row1)
  }
}

colnames(clusts) <- c("Season", "Allocation", "Probability")

clusts
```


```{r}
### quantile bootstrap stuff
jack_boot = function(data, u_val){
  
  DataFit <- gpd.fit(data, u_val, show=FALSE)

  sca <- DataFit$mle[1]
  sha <- DataFit$mle[2]
  
  quant_all_data <- qgpd(0.9999, loc=u_val, scale = sca, shape = sha)
  
  quantiles <- rep(0)
  for (i in 1:length(data)){
    data_to_fit <- data[-i]
    for (j in 1:length(data_to_fit)){
      if (j >= i){
        data_to_fit2 <- data_to_fit[-j]
        DataFit <- gpd.fit(data_to_fit2, u_val, show=FALSE)

        sca <- DataFit$mle[1]
        sha <- DataFit$mle[2]
  
        quantiles[i] <- qgpd(0.9999, loc=u_val, scale = sca, shape = sha)
        
      }
    }
  }
  vals <- quantile(quantiles, probs = c(0.25, 0.75))
  return(c(quant_all_data, quant_all_data-as.numeric(vals[1]), as.numeric(vals[2])-quant_all_data))
}

```


### S1 cluster 1


```{r}

GPD_S1_1_Data <- YsAmaDataS1_nona[cluster1_alocsS1,'Y']

hist(GPD_S1_1_Data)
mrl.plot(GPD_S1_1_Data)

u_0 <- 120



gpd.fit(GPD_S1_1_Data, u_0)

uvec = seq(from = 80, to = 130, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S1_1_Datafit <- gpd.fit(GPD_S1_1_Data, uvec[i])
  
  ModScal[i] <- GPD_S1_1_Datafit$mle[1]
  Shape[i] <- GPD_S1_1_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S1_1_Datafit$se[1]
  ShapeErr[i] <- GPD_S1_1_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 100

ExtremeValues <- GPD_S1_1_Data[GPD_S1_1_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S1_1_DataFit <- gpd.fit(GPD_S1_1_Data, u_0)

sca <- FinalGPD_S1_1_DataFit$mle[1]
sha <- FinalGPD_S1_1_DataFit$mle[2]


Initial_Cov <- FinalGPD_S1_1_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S1_1_DataFit)

```


### Bootstrap


```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust1_s1quant <- quants[1]
clust1_s1lowquant <- quants[2]
clust1_s1highquant <- quants[3]

```



### S1 cluster 2


```{r}

GPD_S1_2_Data <- YsAmaDataS1_nona[cluster2_alocsS1,'Y']
hist(GPD_S1_2_Data)
mrl.plot(GPD_S1_2_Data)

u_0 <- 120



gpd.fit(GPD_S1_2_Data, u_0)

uvec = seq(from = 80, to = 130, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S1_2_Datafit <- gpd.fit(GPD_S1_2_Data, uvec[i])
  
  ModScal[i] <- GPD_S1_2_Datafit$mle[1]
  Shape[i] <- GPD_S1_2_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S1_2_Datafit$se[1]
  ShapeErr[i] <- GPD_S1_2_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 95

ExtremeValues <- GPD_S1_2_Data[GPD_S1_2_Data>u_0]
ExtremeValues

df[df$ID == "a", ]

FinalGPD_S1_2_DataFit <- gpd.fit(GPD_S1_2_Data, u_0)

sca <- FinalGPD_S1_2_DataFit$mle[1]
sha <- FinalGPD_S1_2_DataFit$mle[2]



Initial_Cov <- FinalGPD_S1_2_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10

gpd.diag(FinalGPD_S1_2_DataFit)

```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust2_s1quant <- quants[1]
clust2_s1lowquant <- quants[2]
clust2_s1highquant <- quants[3]

```


### S1 cluster 3


```{r}

GPD_S1_3_Data <- YsAmaDataS1_nona[cluster3_alocsS1,'Y']
mrl.plot(GPD_S1_3_Data)

u_0 <- 120



gpd.fit(GPD_S1_3_Data, u_0)

uvec = seq(from = 80, to = 130, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S1_3_Datafit <- gpd.fit(GPD_S1_3_Data, uvec[i])
  
  ModScal[i] <- GPD_S1_3_Datafit$mle[1]
  Shape[i] <- GPD_S1_3_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S1_3_Datafit$se[1]
  ShapeErr[i] <- GPD_S1_3_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 100

ExtremeValues <- GPD_S1_3_Data[GPD_S1_3_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S1_3_DataFit <- gpd.fit(GPD_S1_3_Data, u_0)

sca <- FinalGPD_S1_3_DataFit$mle[1]
sha <- FinalGPD_S1_3_DataFit$mle[2]


Initial_Cov <- FinalGPD_S1_3_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S1_3_DataFit)

```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust3_s1quant <- quants[1]
clust3_s1lowquant <- quants[2]
clust3_s1highquant <- quants[3]

```



### S1 cluster 4


```{r}

GPD_S1_4_Data <- YsAmaDataS1_nona[cluster4_alocsS1,'Y']
mrl.plot(GPD_S1_4_Data)

u_0 <- 120



gpd.fit(GPD_S1_4_Data, u_0)

uvec = seq(from = 80, to = 120, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S1_4_Datafit <- gpd.fit(GPD_S1_4_Data, uvec[i])
  
  ModScal[i] <- GPD_S1_4_Datafit$mle[1]
  Shape[i] <- GPD_S1_4_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S1_4_Datafit$se[1]
  ShapeErr[i] <- GPD_S1_4_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 100

ExtremeValues <- GPD_S1_4_Data[GPD_S1_4_Data>u_0]


FinalGPD_S1_4_DataFit <- gpd.fit(GPD_S1_4_Data, u_0)

sca <- FinalGPD_S1_4_DataFit$mle[1]
sha <- FinalGPD_S1_4_DataFit$mle[2]


Initial_Cov <- FinalGPD_S1_4_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S1_4_DataFit)

```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust4_s1quant <- quants[1]
clust4_s1lowquant <- quants[2]
clust4_s1highquant <- quants[3]

```



### S1 cluster 5


```{r}

GPD_S1_5_Data <- YsAmaDataS1_nona[cluster5_alocsS1,'Y']
mrl.plot(GPD_S1_5_Data)

u_0 <- 120



gpd.fit(GPD_S1_5_Data, u_0)

uvec = seq(from = 80, to = 130, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S1_5_Datafit <- gpd.fit(GPD_S1_5_Data, uvec[i])
  
  ModScal[i] <- GPD_S1_5_Datafit$mle[1]
  Shape[i] <- GPD_S1_5_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S1_5_Datafit$se[1]
  ShapeErr[i] <- GPD_S1_5_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 95

ExtremeValues <- GPD_S1_5_Data[GPD_S1_5_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S1_5_DataFit <- gpd.fit(GPD_S1_5_Data, u_0)

sca <- FinalGPD_S1_5_DataFit$mle[1]
sha <- FinalGPD_S1_5_DataFit$mle[2]


Initial_Cov <- FinalGPD_S1_5_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))
gpd.diag(FinalGPD_S1_5_DataFit)
```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust5_s1quant <- quants[1]
clust5_s1lowquant <- quants[2]
clust5_s1highquant <- quants[3]

```


### Store vals

```{r}
clust_quants_S1 <- c(clust1_s1quant,clust2_s1quant,clust3_s1quant,clust4_s1quant,clust5_s1quant)
clust_quants_S1low <- c(clust1_s1lowquant,clust2_s1lowquant,clust3_s1lowquant,clust4_s1lowquant,clust5_s1lowquant)
clust_quants_S1high <- c(clust1_s1highquant,clust2_s1highquant,clust3_s1highquant,clust4_s1highquant,clust5_s1highquant)
```



### S2 cluster 1


```{r}

GPD_S2_1_Data <- YsAmaDataS2_nona[cluster1_alocsS2,'Y']

hist(GPD_S2_1_Data)

mrl.plot(GPD_S2_1_Data)


u_0 <- 120


gpd.fit(GPD_S2_1_Data, u_0)

uvec = seq(from = 80, to = 130, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S2_1_Datafit <- gpd.fit(GPD_S2_1_Data, uvec[i])
  
  ModScal[i] <- GPD_S2_1_Datafit$mle[1]
  Shape[i] <- GPD_S2_1_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S2_1_Datafit$se[1]
  ShapeErr[i] <- GPD_S2_1_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 100

ExtremeValues <- GPD_S2_1_Data[GPD_S2_1_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S2_1_DataFit <- gpd.fit(GPD_S2_1_Data, u_0)

sca <- FinalGPD_S2_1_DataFit$mle[1]
sha <- FinalGPD_S2_1_DataFit$mle[2]


Initial_Cov <- FinalGPD_S2_1_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S2_1_DataFit)

```


### Bootstrap Stuff


```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust1_s2quant <- quants[1]
clust1_s2lowquant <- quants[2]
clust1_s2highquant <- quants[3]

```







### S2 cluster 2


```{r}

GPD_S2_2_Data <- YsAmaDataS2_nona[cluster2_alocsS2,'Y']
hist(GPD_S2_2_Data)

mrl.plot(GPD_S2_2_Data)

u_0 <- 120



gpd.fit(GPD_S2_2_Data, u_0)

uvec = seq(from = 80, to = 130, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S2_2_Datafit <- gpd.fit(GPD_S2_2_Data, uvec[i])
  
  ModScal[i] <- GPD_S2_2_Datafit$mle[1]
  Shape[i] <- GPD_S2_2_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S2_2_Datafit$se[1]
  ShapeErr[i] <- GPD_S2_2_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 100

ExtremeValues <- GPD_S2_2_Data[GPD_S2_2_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S2_2_DataFit <- gpd.fit(GPD_S2_2_Data, u_0)

sca <- FinalGPD_S2_2_DataFit$mle[1]
sha <- FinalGPD_S2_2_DataFit$mle[2]



Initial_Cov <- FinalGPD_S2_2_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S2_2_DataFit)

```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust2_s2quant <- quants[1]
clust2_s2lowquant <- quants[2]
clust2_s2highquant <- quants[3]

```


### S2 cluster 3


```{r}

GPD_S2_3_Data <- YsAmaDataS2_nona[cluster3_alocsS2,'Y']

mrl.plot(GPD_S2_3_Data)


u_0 <- 110



gpd.fit(GPD_S2_3_Data, u_0)

uvec = seq(from = 80, to = 110, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S2_3_Datafit <- gpd.fit(GPD_S2_3_Data, uvec[i])
  
  ModScal[i] <- GPD_S2_3_Datafit$mle[1]
  Shape[i] <- GPD_S2_3_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S2_3_Datafit$se[1]
  ShapeErr[i] <- GPD_S2_3_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 80

ExtremeValues <- GPD_S2_3_Data[GPD_S2_3_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S2_3_DataFit <- gpd.fit(GPD_S2_3_Data, u_0)

sca <- FinalGPD_S2_3_DataFit$mle[1]
sha <- FinalGPD_S2_3_DataFit$mle[2]


Initial_Cov <- FinalGPD_S2_3_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S2_3_DataFit)

```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust3_s2quant <- quants[1]
clust3_s2lowquant <- quants[2]
clust3_s2highquant <- quants[3]

```



### S2 cluster 4


```{r}

GPD_S2_4_Data <- YsAmaDataS2_nona[cluster4_alocsS2,'Y']
mrl.plot(GPD_S2_4_Data)

u_0 <- 120



gpd.fit(GPD_S2_4_Data, u_0)

uvec = seq(from = 80, to = 120, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S2_4_Datafit <- gpd.fit(GPD_S2_4_Data, uvec[i])
  
  ModScal[i] <- GPD_S2_4_Datafit$mle[1]
  Shape[i] <- GPD_S2_4_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S2_4_Datafit$se[1]
  ShapeErr[i] <- GPD_S2_4_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 85

ExtremeValues <- GPD_S2_4_Data[GPD_S2_4_Data>u_0]


FinalGPD_S2_4_DataFit <- gpd.fit(GPD_S2_4_Data, u_0)

sca <- FinalGPD_S2_4_DataFit$mle[1]
sha <- FinalGPD_S2_4_DataFit$mle[2]


Initial_Cov <- FinalGPD_S2_4_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))

gpd.diag(FinalGPD_S2_4_DataFit)

```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust4_s2quant <- quants[1]
clust4_s2lowquant <- quants[2]
clust4_s2highquant <- quants[3]

```



### S2 cluster 5


```{r}

GPD_S2_5_Data <- YsAmaDataS2_nona[cluster5_alocsS2,'Y']
mrl.plot(GPD_S2_5_Data)

u_0 <- 110



gpd.fit(GPD_S2_5_Data, u_0)

uvec = seq(from = 80, to = 110, by = 5)
uvec

ModScal <- vector(length = length(uvec))
Shape <- vector(length = length(uvec))

ModScalErr <- vector(length = length(uvec))
ShapeErr <- vector(length = length(uvec))

for(i in 1:length(uvec)){
  GPD_S2_5_Datafit <- gpd.fit(GPD_S2_5_Data, uvec[i])
  
  ModScal[i] <- GPD_S2_5_Datafit$mle[1]
  Shape[i] <- GPD_S2_5_Datafit$mle[2]
  
  ModScalErr[i] <- GPD_S2_5_Datafit$se[1]
  ShapeErr[i] <- GPD_S2_5_Datafit$se[2]
}


df <- data.frame(uvec, ModScal, ModScalErr, Shape, ShapeErr)

p<- ggplot(df, aes(x=uvec, y=ModScal)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=ModScal-ModScalErr, ymax=ModScal+ModScalErr), width=.4)
p<-p+labs(title="Parameter Estimate of Modified Scale", x="Threshold", y = "Modified Scale")


p2<- ggplot(df, aes(x=uvec, y=Shape)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=Shape-ShapeErr, ymax=Shape+ShapeErr), width=.4)
p2<-p2+labs(title="Parameter Estimate of Shape", x="Threshold", y = "Shape")

ggarrange(p, p2, nrow = 2, labels = c("A", "B")) 
```


```{r}

u_0 <- 80

ExtremeValues <- GPD_S2_5_Data[GPD_S2_5_Data>u_0]

df[df$ID == "a", ]

FinalGPD_S2_5_DataFit <- gpd.fit(GPD_S2_5_Data, u_0)

sca <- FinalGPD_S2_5_DataFit$mle[1]
sha <- FinalGPD_S2_5_DataFit$mle[2]


Initial_Cov <- FinalGPD_S2_5_DataFit$cov


CI_max <- sha + 1.96*sqrt(Initial_Cov[2,2])
CI_min <- sha - 1.96*sqrt(Initial_Cov[2,2])

```

```{r}
gpd_dens = function(x, xi, sig){
  return(1/sig*(1+xi*x/sig)^(-1/xi-1))
}
gpd_x = c(0:200)/10
plot(gpd_x,gpd_dens(gpd_x, sha, sca))
gpd.diag(FinalGPD_S2_5_DataFit)
```

```{r}

quants <- jack_boot(ExtremeValues, u_0)
clust5_s2quant <- quants[1]
clust5_s2lowquant <- quants[2]
clust5_s2highquant <- quants[3]

```


### Store vals

```{r}
clust_quants_S2 <- c(clust1_s2quant,clust2_s2quant,clust3_s2quant,clust4_s2quant,clust5_s2quant)
clust_quants_S2low <- c(clust1_s2lowquant,clust2_s2lowquant,clust3_s2lowquant,clust4_s2lowquant,clust5_s2lowquant)
clust_quants_S2high <- c(clust1_s2highquant,clust2_s2highquant,clust3_s2highquant,clust4_s2highquant,clust5_s2highquant)
```

### Final Quantile Allocations (mixture over clusters; Challenge 1 paper)

```{r}

target_tail <- 1e-4  # P(Y > q | X=x) = 0.0001  <=>  P(Y <= q | X=x) = 0.9999

# Collect per-cluster GPD fits and raw Y data (needed to estimate zeta_u empirically)
fits_S1 <- list(FinalGPD_S1_1_DataFit, FinalGPD_S1_2_DataFit, FinalGPD_S1_3_DataFit,
                FinalGPD_S1_4_DataFit, FinalGPD_S1_5_DataFit)
data_S1 <- list(GPD_S1_1_Data, GPD_S1_2_Data, GPD_S1_3_Data, GPD_S1_4_Data, GPD_S1_5_Data)

fits_S2 <- list(FinalGPD_S2_1_DataFit, FinalGPD_S2_2_DataFit, FinalGPD_S2_3_DataFit,
                FinalGPD_S2_4_DataFit, FinalGPD_S2_5_DataFit)
data_S2 <- list(GPD_S2_1_Data, GPD_S2_2_Data, GPD_S2_3_Data, GPD_S2_4_Data, GPD_S2_5_Data)

# Tail probability estimate within a cluster:
# P(Y > q | Z=j) ≈ zeta_u_j * P(Y > q | Y > u_j, Z=j) using a fitted GPD above u_j
cluster_tailprob <- function(q, fit, y){
  u <- fit$threshold
  sig <- fit$mle[1]
  xi  <- fit$mle[2]
  zeta_u <- mean(y > u)

  if (q <= u) return(zeta_u)  # q will be extreme, so usually q > u

  if (abs(xi) < 1e-10){
    # xi -> 0 limit
    return(zeta_u * exp(-(q - u) / sig))
  } else {
    inner <- 1 + xi * (q - u) / sig
    if (inner <= 0) return(0)
    return(zeta_u * inner^(-1/xi))
  }
}

# Solve for q such that sum_j w_j * P(Y > q | Z=j) = target_tail
solve_mixture_quantile <- function(weights, fits, ys, target_tail){
  uvec <- sapply(fits, function(f) f$threshold)
  lower <- max(uvec) + 1e-6

  f <- function(q){
    sum(sapply(1:length(fits), function(j) weights[j] * cluster_tailprob(q, fits[[j]], ys[[j]]))) - target_tail
  }

  # Find an upper bracket where f(upper) < 0
  upper <- lower + 1
  it <- 0
  while (f(upper) > 0 && it < 60){
    upper <- upper * 1.5
    it <- it + 1
  }
  if (it == 60) upper <- upper + 1e4  # fallback

  uniroot(f, lower = lower, upper = upper)$root
}

# Simple parametric uncertainty: simulate (sigma, xi) from asymptotic normal approx
# and compute empirical 25%/75% of the resulting mixture quantiles.
mixture_quantile_ci <- function(weights, fits, ys, target_tail, B = 200){
  # point estimate
  q_hat <- solve_mixture_quantile(weights, fits, ys, target_tail)

  q_sims <- rep(NA_real_, B)
  for (b in 1:B){
    sim_fits <- fits
    for (j in 1:length(fits)){
      mu <- fits[[j]]$mle
      V  <- fits[[j]]$cov
      draw <- as.numeric(rmvnorm(1, mean = mu, sigma = V))

      # enforce sigma > 0
      tries <- 0
      while (draw[1] <= 0 && tries < 50){
        draw <- as.numeric(rmvnorm(1, mean = mu, sigma = V))
        tries <- tries + 1
      }
      if (draw[1] <= 0) draw[1] <- mu[1]  # fallback

      sim_fits[[j]]$mle <- draw
    }

    q_sims[b] <- tryCatch(
      solve_mixture_quantile(weights, sim_fits, ys, target_tail),
      error = function(e) NA_real_
    )
  }

  q_sims <- q_sims[is.finite(q_sims)]
  if (length(q_sims) < 10){
    return(c(q_hat, NA_real_, NA_real_))
  }

  qs <- quantile(q_sims, probs = c(0.25, 0.75), na.rm = TRUE)
  c(q_hat, as.numeric(qs[1]), as.numeric(qs[2]))
}

finalquantiles <- data.frame(Quantile = numeric(), Low = numeric(), High = numeric())

for (i in 1:nrow(clusts)){
  season <- clusts[i, "Season"]
  w <- as.numeric(clusts[i, paste0("p", 1:5)])
  w <- w / sum(w)

  if (season == "S1"){
    res <- mixture_quantile_ci(w, fits_S1, data_S1, target_tail, B = 200)
  } else {
    res <- mixture_quantile_ci(w, fits_S2, data_S2, target_tail, B = 200)
  }

  finalquantiles <- rbind(finalquantiles, data.frame(Quantile = res[1], Low = res[2], High = res[3]))
}

finalquantiles


```


```{r}
save(finalquantiles, file = "AnswerC1.RData")

```



